\chapter*{Acknowledgements}

I have always had in mind the idea of pursuing a Ph.D. program, as I have always liked learning. However, after finishing my bachelor degree, I found great professional opportunities that I could not overlook. Nonetheless, I never lost the desire to continue my educational journey.  Thus, a  few years later, I found the fantastic opportunity to enrol into a part-time Ph.D. at \textit{Universitat de València}. Being a part-time Ph.D. student at the \textit{Intelligent Data Analysis Laboratory} (IDAL) of the \textit{Universitat de València} was lots of fun, and definitely this was one of the best decisions I could have ever made.  

I would like to start by sincerely expressing my gratitude to \textit{Emilio Soria-Olivas, PhD}, my advisor. First of all, thank you for instilling your passion in artificial intelligence into me. Your encouragement and guidance has been a key factor in the development of this thesis, without which this work could never have been possible. I would also like to thank all the members of the IDAL, many of whom directly collaborated and reviewed my work: \textit{Antonio-José Serrano-López, Ph.D., Marcelino  Martínez-Sober, Ph.D., Juan Gómez-Sanchis, Ph.D., Fernando Mateo-Jiménez, Ph.D and Joan Vila-Francés, Ph.D}. I hope to have the opportunity to collaborate with you in future projects.

On the professional side, I would like to thank all the \textit{Alexa TTS} research group, for accepting me into their family. Thanks to all the people who directly collaborated on my contributions to the speech generation field: \textit{Roberto Barra-Chicote, Ph.D., Jasha Droppo, Ph.D., Julian Roth, M.Sc. and Grzegorz Beringer, M.Sc}. Special mention to \textit{Roberto Barra-Chicote, Ph.D.}, my mentor at \textit{Alexa}, for all his guidance and support, and for continuously boosting my motivation, helping me be a better scientist.

Finally, and more importantly, I would like to acknowledge my family, especially my partner \textit{María}, my brother \textit{David} and my parents \textit{Reme} and \textit{Ignacio}, for all the unconditional support during these years, in all its modalities. This work would not have been possible without your love.

I hope you, the reader, enjoy this thesis as much as I enjoyed writing it. 
\clearpage
